---
title: "Winter Wheat (2023) Fertilization Timing Analysis"
author: "Ahmed Tohameem Alam"
date: "`r Sys.Date()`"
email: "ahmedtohameem@gmail.com"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## Introduction

This document analyzes data from the "StaPraxRegio" Database, focusing on weather, development stages, and attempts_tab. The objectives are to clean, merge, and conduct exploratory data analysis. Visualizations will depict the differences between targeted and actual fertilization dates and the distribution of fertilization dates at specific stages across various regions in Germany.

## Load Required Libraries

I start by loading the necessary R libraries.

```{r load-libraries, message=FALSE, warning=FALSE}
rm(list = ls())
setwd("D:/OneDrive - b-tu.de/Msc_Project/codes_TM/Results/WW_2023")
library(readxl) # read excel files into R.
library(writexl) # write data frames to excel files.
library(dplyr) # Includes tools for data manipulation, including filtering, summarizing and transforming data.
library(lubridate) # Simplifies working with dates and times in R.
library(knitr) # Facilitates dynamic report generation, combining R code and narrative. 
library(ggplot2) # Creates elegant data visualizations using the grammar of graphics.
library(tidyr) # Helps tidy data by reshaping it into a consistent format. 
library(stringr) # Provides functions for string manipulations and regular expressions. 
library(patchwork) # Combines multiple ggplot2 plots into a single layout. 
library(DT) # Creates interactive, dynamic tables in R, typically used in web applications.
library(readr) # provides functions for reading and writing data files
library(stringr) # offers tools for string manipulation and regular expression handling.
library(ggpattern) #extension of ggplot2 that enables the addition of fill patterns to geoms, enhancing the visualization of categorical data.
library(broom) # used to convert statistical analysis objects into tidy data frames, making it easier to work with and visualize the results.
```

## 1. Define the file path and filter data

```{r}
file_path <- "D:/OneDrive - b-tu.de/Msc_Project/StaPrax Regio Database/Final DB/DB_2023_v2.xlsx"
```

### Read the specific sheets

```{r}
weather_data <- read_excel(file_path, sheet = "Weather_Attempt_10d")
development_data <- read_excel(file_path, sheet = "Development Stages")
attempts_data <- read_excel(file_path, sheet = "Attempts")
analysis <- read_csv("D:/OneDrive - b-tu.de/Msc_Project/StaPrax Regio Database/Final DB/Analysis.csv")
```

### Filter attepmts_tab and analysis file to keep only winter wheat and year 2023. 

```{r}
attempts_data <- attempts_data %>%
  filter(Crop == "WW")
# this tab contains the actual fertilization dates. I am filtering it for the crop winter wheat and for one year : 2023
attempts_data <- attempts_data %>%
  filter(Year == "2023")
```
```{r}
#This analysis file contains information about field trials that are unsuitable for research as they were compromised. 
analysis <- analysis %>%
  filter(CropType_Long == "Winter Soft Wheat")
analysis <- analysis %>%
  filter(Year == "2023")
length(unique(analysis$Attempt))
field <- unique(analysis$Attempt) # A check to see how many unique Attempt ID we have in current analysis.
```


## 2. Further filtering of the tabs "development_stages" & "attempts"

### Use left_join to retain rows from attempts_data and match where possible

```{r}
attempts_data_with_ID_WS <- attempts_data %>%
  left_join(weather_data %>% select(ID_WS, Attempt), by = "Attempt")
#using the left join function I am bringing the ID_WS column to the attempts_tab, which will later help merge it with the developments_tab
```

```{r paged.print=TRUE}
# Rename 'Year' to 'Period' in attempts_data_with_ID_WS, as it is named as Period in the developments_tab
attempts_data_with_ID_WS <- attempts_data_with_ID_WS %>%
  rename(Period = Year)
# Some experiments need to be excluded as they contain errors from the field
attempts_data_with_ID_WS <- attempts_data_with_ID_WS %>%
  filter(Attempt %in% field)
# This step is done so that compromised trials/attempts get excluded. 
```


### Select only the required columns in the 'development_data' dataframe

```{r}
development_data <- development_data[, c("ID", "ID_WS", "Period", "VB_GL_korrTSum", "BBCH32_mod", "BBCH37_mod", "BBCH49_mod")]
```

### Filter rows to keep only those with "2023" in the 'Period' column

```{r}
development_data <- development_data[development_data$Period %in% c("2023"),]

```

### Keep only rows in 'development_data' that have matching 'ID_WS' in 'weather_data'

```{r}
development_data <- development_data %>%
  semi_join(weather_data, by = "ID_WS")
# Because development_tab contains a lot of time periods but my focus is only on winter soft wheat and the year 2023
# So I do not need all the rows from this tab, only the ones that match this analysis requirements. 
```

## 3. Merging attempts_tab and developments_tab

### Perform the join using ID_WS

```{r paged.print=TRUE}
# Step 1: Merge 'attempts_data_with_ID_WS' with 'development_data' based on 'ID_WS'
combined_data <- attempts_data_with_ID_WS %>%
  left_join(development_data, by = "ID_WS")

# Step 2: Create a new column 'modeled_T1' in 'combined_data'
combined_data <- combined_data %>%
  mutate(modeled_T1 = VB_GL_korrTSum - 15)

# Step 3: Select only the required columns for analysis
selected_data <- combined_data %>%
  select(
    ID, ID_WS, GPS_noerdlB, VB_GL_korrTSum, modeled_T1, BBCH32_mod, BBCH37_mod, BBCH49_mod, 
    Attempt, Place, Crop, Soil_identification, ID_Nature, BKR_ID, 
    Nmin_0_30, Nmin_30_60, Nmin_60_90, Nmin_DBE, Date_sowing, N_Level, 
    Date_T1, Date_T2, Date_T3, Date_T4, Date_T5, Date_harvest
  )
 

```




## 4. Missing value analysis

```{r}

# Count rows with any missing values in combined_data
rows_with_missing <- selected_data %>%
  filter(if_any(everything(), is.na)) %>%
  nrow()
# View the rows with missing values
rows_with_missing_data <- selected_data %>%
  filter(if_any(everything(), is.na))

# Print the rows with missing values
datatable(rows_with_missing_data, options = list(pagelength = 10, autowidth= T))


# Only 4 rows have some values missing. Relatively low amount.  
```

### Check for duplicates based on 'ID_WS' columns

```{r}
# Identify all rows with duplicate ID_WS
duplicates_by_ID_WS <- selected_data %>%
  group_by(ID_WS) %>%
  filter(n() > 1) %>%
  ungroup()  # Ungroup for further operations if needed

datatable(duplicates_by_ID_WS, options = list(pagelength = 10, autowidth= T))
# We have few Attempt with same ID_WS but they have different Attempt ID meaning they are not exactly duplicates.  
```


# 5. Fertilization date deviation check of Winter wheat

### Converting to DOY

```{r}
# Convert the date columns to day of the year (1-365)
filtered_data <- selected_data %>%
  mutate(Day_sowing = yday(as.Date(Date_sowing)),
         Day_T1 = yday(as.Date(Date_T1)),
         Day_T2 = yday(as.Date(Date_T2)),
         Day_T3 = yday(as.Date(Date_T3)),
         Day_T4 = yday(as.Date(Date_T4)),
         Day_T5 = yday(as.Date(Date_T5)),
         Day_harvest = yday(as.Date(Date_harvest)))
# This conversion is necessary to calculate the deviation of days
```


### Calculate the differences between day columns and fractional columns (modelled DOY's have fractional values)

```{r}


# Create a new column to store median of actual T4 timings
filtered_data <- filtered_data %>%
  mutate(medianT4 = median(Day_T4, na.rm = TRUE))
# Modeled T4 dates vary significantly so median is being used for T4 fertilization stage.   
# Calculate differences and add them as new columns
filtered_data <- filtered_data %>%
  mutate(
         Diff_T1_model  = Day_T1 - modeled_T1,
         Diff_T1_T2     = Day_T2 - Day_T1, # Difference between Actual T2 and Actual T1 (should be at least 7)
         Diff_T2_VB_GL  = Day_T2 - VB_GL_korrTSum,
         Diff_T3_BBCH32 = Day_T3 - BBCH32_mod,
         Diff_T4_BBCH37 = Day_T4 - medianT4, # Using median for Targeted T4, as most modeled dates fail here.
         Diff_T5_BBCH49 = Day_T5 - BBCH49_mod)
```


### Check the new filter data

```{r}
datatable(filtered_data, options = list(pagelength = 10, autowidth = T))

```


### Deviance of the two fertilization dates

```{r}

# Gather difference columns into long format for easier plotting
plot_data <- filtered_data %>%
  select(Attempt, Diff_T2_VB_GL, Diff_T1_T2, Diff_T1_model, Diff_T3_BBCH32, Diff_T4_BBCH37, Diff_T5_BBCH49) %>%
  pivot_longer(cols = c(Diff_T2_VB_GL, Diff_T1_T2, Diff_T1_model, Diff_T3_BBCH32, Diff_T4_BBCH37, Diff_T5_BBCH49),
               names_to = "Difference_Variable", values_to = "Difference_Value") %>%
  drop_na()

# Create color categories based on specified day ranges
plot_data <- plot_data %>%
  mutate(
    Color_Category = case_when(
      abs(Difference_Value) <= 10 ~ "Deviation below or equal to 10 days",   # Grey
      abs(Difference_Value) >= 11 & abs(Difference_Value) <= 19 ~ "Deviation between 11 and 19 days",  # Pink
      abs(Difference_Value) >= 20 ~ "Deviation of 20 days or more"  # Dark Red
    )
  )

# Wrap long Attempt labels (e.g., after every 10 characters)
plot_data$Attempt <- str_wrap(plot_data$Attempt, width = 10)

# Function to create plots for each difference variable with specified color categories
create_plot <- function(data, variable, title) {
  ggplot(data %>% filter(Difference_Variable == variable), 
         aes(x = Attempt, y = Difference_Value, fill = Color_Category)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_manual(values = c(
      "Deviation below or equal to 10 days" = "gray",
      "Deviation between 11 and 19 days" = "pink",
      "Deviation of 20 days or more" = "darkred"
    )) + 
    theme_minimal() +
    labs(title = title,
         x = "Attempts",
         y = "Difference Value") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8))
}

# List of variables and titles for the plots
variables <- c("Diff_T2_VB_GL", "Diff_T3_BBCH32", "Diff_T4_BBCH37", "Diff_T5_BBCH49")
titles <- c("Fertilization Date Deviations: Difference of Targetted & actual T2",
            "Fertilization Date Deviations: Difference of Targetted & actual T3",
            "Fertilization Date Deviations: Difference of Targetted & actual T4",
            "Fertilization Date Deviations: Difference of Targetted & actual T5"
            )

# Generate and store each plot in a list
plots <- lapply(seq_along(variables), function(i) {
  create_plot(plot_data, variables[i], titles[i])
})

# Print all plots
for (plot in plots) {
  print(plot)
}
```


### Plotting Modelled T1 and Actual T1 timings
```{r}


# Prepare the data
plot_data_t1 <- filtered_data %>%
  filter(!is.na(Diff_T1_model) & !is.na(Diff_T1_T2)) %>%
  select(Attempt, Diff_T1_model, Diff_T1_T2) %>%
  mutate(
    Color_Category = case_when(
      Diff_T1_model > 19 ~ "Deviation of more than 19 days",        
      Diff_T1_model >= 10 & Diff_T1_model <= 19 ~ "Deviation of +10 to +19 days",  
      TRUE ~ "Other deviations"                                     
    ),
    Pattern = if_else(Diff_T1_T2 < 7, "Difference between Actual T2 & Actual T1 < 7d", "Difference between Actual T2 & Actual T1 > 7d")
  )

# Wrap long Attempt labels
plot_data_t1$Attempt <- str_wrap(plot_data_t1$Attempt, width = 10)

# Create the corrected plot
ggplot(plot_data_t1, aes(x = Attempt, y = Diff_T1_model)) +
  geom_bar_pattern(
    aes(fill = Color_Category, pattern = Pattern),
    stat = "identity", position = "dodge",
    pattern_fill = "black", pattern_density = 0.1, pattern_spacing = 0.05
  ) +
  scale_fill_manual(
    name = "Deviation Category",
    values = c(
      "Deviation of more than 19 days" = "darkred",
      "Deviation of +10 to +19 days" = "pink",
      "Other deviations" = "gray"
    )
  ) +
  scale_pattern_manual(
    name = "Pattern",
    values = c(
      "Difference between Actual T2 & Actual T1 < 7d" = "stripe",
      "Difference between Actual T2 & Actual T1 > 7d" = "none"
    )
  ) +
  guides(
    fill = guide_legend(override.aes = list(pattern = "none")),  # Ensures colors appear solid
    pattern = guide_legend(order = 2)  # Keeps pattern legend separate
  ) +
  theme_minimal() +
  labs(
    title = "Fertilization Date Deviations: Targeted & Actual T1",
    x = "Attempts",
    y = "Difference Value for T1"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8))

```


## Spatial Distribuition of fertilization 

### Scatter Plot with Date_sowing

```{r}

# Step 1: Ensure Date columns are in Date format and set specific year for model dates
selected_data <- selected_data %>%
  mutate(
    Date_T1 = as.Date(Date_T1, format = "%Y-%m-%d"),
    Date_T2 = as.Date(Date_T2, format = "%Y-%m-%d"),
    Date_T3 = as.Date(Date_T3, format = "%Y-%m-%d"),
    Date_T4 = as.Date(Date_T4, format = "%Y-%m-%d"),
    Date_T5 = as.Date(Date_T5, format = "%Y-%m-%d"),
    Model_Date_T1 = as.Date(modeled_T1, format = "%Y-%m-%d"),
    Model_Date_T2 = as.Date(VB_GL_korrTSum, format = "%Y-%m-%d"),
    Model_Date_T3 = as.Date(BBCH32_mod, format = "%Y-%m-%d"),
    Model_Date_T4 = as.Date(BBCH37_mod, format = "%Y-%m-%d"),
    Model_Date_T5 = as.Date(BBCH49_mod, format = "%Y-%m-%d"),
    Date_sowing = as.Date(Date_sowing, format = "%Y-%m-%d")
  ) %>%
  mutate(across(starts_with("Model_Date"), ~ `year<-`(.x, 2023)))

# Step 2: Drop NA values for relevant columns
selected_data <- selected_data %>%
  drop_na(Date_T1, Date_T2, Date_T3, Date_T4, Date_T5, Place, GPS_noerdlB, Model_Date_T1, Model_Date_T2, Model_Date_T3, Model_Date_T4, Model_Date_T5, Date_sowing)

# Step 3: Arrange dataframe by increasing GPS_noerdlB (latitude)
selected_data <- selected_data %>%
  arrange(GPS_noerdlB)

# Step 4: Gather both actual and model dates into long format for plotting, excluding Date_sowing
spatial_plot_data <- selected_data %>%
  select(Place, GPS_noerdlB, Date_T1, Date_T2, Date_T3, Date_T4, Date_T5,
         Model_Date_T1, Model_Date_T2, Model_Date_T3, Model_Date_T4, Model_Date_T5) %>%
  pivot_longer(cols = c(Date_T1, Date_T2, Date_T3, Date_T4, Date_T5, 
                        Model_Date_T1, Model_Date_T2, Model_Date_T3, Model_Date_T4, Model_Date_T5),
               names_to = "Fertilization_Stage", values_to = "Date_Value") %>%
  mutate(
    Type = ifelse(str_detect(Fertilization_Stage, "Model_"), "Model", "Actual"),  # Mark as Model or Actual
    Fertilization_Stage = str_replace(Fertilization_Stage, "Model_", "")  # Normalize stage names for grouping
  )

# Step 5: Convert Place to factor ordered by latitude for consistent plotting
spatial_plot_data <- spatial_plot_data %>%
  mutate(Place = factor(Place, levels = unique(selected_data$Place[order(selected_data$GPS_noerdlB)])))

# Step 6: Get unique fertilization stages
fertilization_stages <- unique(spatial_plot_data$Fertilization_Stage)

# Step 7: Loop through each fertilization stage and plot with ordered Place by latitude
for (stage in fertilization_stages) {
  # Filter data for the current fertilization stage
  stage_data <- spatial_plot_data %>% filter(Fertilization_Stage == stage)
  
  # Plot for the current stage with Actual, Model, and Sowing dates
  plot <- ggplot(stage_data, aes(x = Place, y = Date_Value, color = Type, shape = Type)) +
    geom_point(size = 3) +
    geom_point(data = selected_data, aes(x = Place, y = Date_sowing, color = "Sowing", shape = "Sowing"), size = 3) +  # Add Date_sowing with specific color and shape
    theme_minimal() +
    labs(title = paste("Fertilization Timing -", stage, "by Place"),
         x = "Place",
         y = "Date") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_color_manual(values = c("Actual" = "black", "Model" = "blue", "Sowing" = "orange")) +  # Include Sowing in color scale
    scale_shape_manual(values = c("Actual" = 16, "Model" = 17, "Sowing" = 18))  # Include Sowing in shape scale
  
  # Print plot
  print(plot)
}



```


### Scatter plot without Date_sowing

```{r}

# Step 5: Convert Place to factor ordered by latitude for consistent plotting
spatial_plot_data <- spatial_plot_data %>%
  mutate(Place = factor(Place, levels = unique(selected_data$Place[order(selected_data$GPS_noerdlB)])))

# Step 6: Get unique fertilization stages
fertilization_stages <- unique(spatial_plot_data$Fertilization_Stage)

# Step 7: Loop through each fertilization stage and plot with ordered Place by latitude
for (stage in fertilization_stages) {
  # Filter data for the current stage
  stage_data <- spatial_plot_data %>% filter(Fertilization_Stage == stage)
  
  # Create plot for the current stage with Actual and Model dates
  plot <- ggplot(stage_data, aes(x = Place, y = Date_Value, color = Type, shape = Type)) +
    geom_point(size = 3) +
    theme_minimal() +
    labs(title = paste("Fertilization Timing -", stage, "by Place"),
         x = "Place",
         y = "Fertilization Date") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_color_manual(values = c("Actual" = "black", "Model" = "blue")) +  # Different colors for Actual and Model
    scale_shape_manual(values = c("Actual" = 16, "Model" = 17))  # Different shapes for Actual and Model
  
  # Print plot
  print(plot)
}


```


### Fertilization timings for different latitudes (Places in Germany)
```{r}

# Step 1: Drop NA and arrange by latitude
latitude_plot_data <- filtered_data %>%
  drop_na(Day_sowing, Day_T1, Day_T2, Day_T3, Day_T4, Day_T5, Day_harvest, Place, GPS_noerdlB, VB_GL_korrTSum, modeled_T1, BBCH32_mod, BBCH37_mod, BBCH49_mod) %>%
  arrange(GPS_noerdlB)

# Step 2: Rename columns for clarity
latitude_plot_data <- latitude_plot_data %>%
  rename(
    modeled_T2 = VB_GL_korrTSum,
    modeled_T3 = BBCH32_mod,
    modeled_T4 = BBCH37_mod,
    modeled_T5 = BBCH49_mod
  )

# Step 3: Create a comparison-ready dataset with pairs of actual and modeled columns
comparison_data <- latitude_plot_data %>%
  select(Place, GPS_noerdlB, Day_T1, modeled_T1, Day_T2, modeled_T2, Day_T3, modeled_T3, 
         Day_T4, modeled_T4, Day_T5, modeled_T5) %>%
  pivot_longer(
    cols = c(Day_T1, modeled_T1, Day_T2, modeled_T2, Day_T3, modeled_T3, Day_T4, modeled_T4, Day_T5, modeled_T5),
    names_to = "Stage_Type", 
    values_to = "Day_Value"
  ) %>%
  mutate(
    Fertilization_Stage = case_when(
      str_detect(Stage_Type, "T1") ~ "T1",
      str_detect(Stage_Type, "T2") ~ "T2",
      str_detect(Stage_Type, "T3") ~ "T3",
      str_detect(Stage_Type, "T4") ~ "T4",
      str_detect(Stage_Type, "T5") ~ "T5"
    ),
    Type = ifelse(str_detect(Stage_Type, "modeled"), "Modeled", "Actual")
  )

# Step 4: Get unique fertilization stages
fertilization_stages <- unique(comparison_data$Fertilization_Stage)

# Step 5: Loop through each fertilization stage and create comparison plots
for (stage in fertilization_stages) {
  # Filter data for the current stage
  stage_data <- comparison_data %>% filter(Fertilization_Stage == stage)
  
  # Create comparison plot for the current stage
  plot <- ggplot(stage_data, aes(x = GPS_noerdlB, y = Day_Value, color = Type, shape = Type)) +
    geom_point(size = 3) +
    # Connect points only within the same Type group
    geom_line(aes(group = interaction(Place, Type)), linetype = "dashed", alpha = 0.5) +
    theme_minimal() +
    labs(title = paste("Comparison of Actual and Modeled Fertilisation Timing -", stage),
         x = "Latitude",
         y = "Fertilization Date") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    scale_x_continuous(
      limits = c(48, max(stage_data$GPS_noerdlB, na.rm = TRUE)),
      breaks = seq(48, max(stage_data$GPS_noerdlB, na.rm = TRUE), by = 0.5)
    ) +
    scale_color_manual(values = c("Actual" = "black", "Modeled" = "blue")) +  # Color-coded comparison
    scale_shape_manual(values = c("Actual" = 16, "Modeled" = 17))  # Shape-coded comparison
  
  # Print plot
  print(plot)
}


```


### Fertilization timing distribution summary statistics
```{r}


# Calculate mean
mean_stats <- latitude_plot_data %>%
  summarise(across(c(Day_sowing, Day_T1, Day_T2, Day_T3, Day_T4, Day_T5, 
                     Day_harvest, GPS_noerdlB, modeled_T1, modeled_T2, 
                     modeled_T3, modeled_T4, modeled_T5), 
                   ~mean(., na.rm = TRUE)))

# Calculate median
median_stats <- latitude_plot_data %>%
  summarise(across(c(Day_sowing, Day_T1, Day_T2, Day_T3, Day_T4, Day_T5, 
                     Day_harvest, GPS_noerdlB, modeled_T1, modeled_T2, 
                     modeled_T3, modeled_T4, modeled_T5), 
                   ~median(., na.rm = TRUE)))

# Calculate minimum
min_stats <- latitude_plot_data %>%
  summarise(across(c(Day_sowing, Day_T1, Day_T2, Day_T3, Day_T4, Day_T5, 
                     Day_harvest, GPS_noerdlB, modeled_T1, modeled_T2, 
                     modeled_T3, modeled_T4, modeled_T5), 
                   ~min(., na.rm = TRUE)))

# Calculate maximum
max_stats <- latitude_plot_data %>%
  summarise(across(c(Day_sowing, Day_T1, Day_T2, Day_T3, Day_T4, Day_T5, 
                     Day_harvest, GPS_noerdlB, modeled_T1, modeled_T2, 
                     modeled_T3, modeled_T4, modeled_T5), 
                   ~max(., na.rm = TRUE)))

# Calculate standard deviation
sd_stats <- latitude_plot_data %>%
  summarise(across(c(Day_sowing, Day_T1, Day_T2, Day_T3, Day_T4, Day_T5, 
                     Day_harvest, GPS_noerdlB, modeled_T1, modeled_T2, 
                     modeled_T3, modeled_T4, modeled_T5), 
                   ~sd(., na.rm = TRUE)))

# Combine the summary statistics into a single data frame
summary_stats <- bind_rows(
  mean_stats %>% mutate(Statistic = "Mean"),
  median_stats %>% mutate(Statistic = "Median"),
  min_stats %>% mutate(Statistic = "Min"),
  max_stats %>% mutate(Statistic = "Max"),
  sd_stats %>% mutate(Statistic = "SD")
) %>% pivot_longer(-Statistic, names_to = "Variable", values_to = "Value")

# Display the summary statistics using `DT::datatable`
datatable(summary_stats, caption = "Summary Statistics for Individual Fertilisation Stage")

```


### Regression analysis for latitude
```{r}

# Linear regression for Day_T1
model_T1 <- lm(Day_T1 ~ GPS_noerdlB, data = latitude_plot_data)
summary_T1 <- tidy(model_T1)
glance_T1 <- glance(model_T1)

# Linear regression for Day_T2
model_T2 <- lm(Day_T2 ~ GPS_noerdlB, data = latitude_plot_data)
summary_T2 <- tidy(model_T2)
glance_T2 <- glance(model_T2)

# Linear regression for Day_T3
model_T3 <- lm(Day_T3 ~ GPS_noerdlB, data = latitude_plot_data)
summary_T3 <- tidy(model_T3)
glance_T3 <- glance(model_T3)

# Linear regression for Day_T4
model_T4 <- lm(Day_T4 ~ GPS_noerdlB, data = latitude_plot_data)
summary_T4 <- tidy(model_T4)
glance_T4 <- glance(model_T4)

# Linear regression for Day_T5
model_T5 <- lm(Day_T5 ~ GPS_noerdlB, data = latitude_plot_data)
summary_T5 <- tidy(model_T5)
glance_T5 <- glance(model_T5)

# Combine necessary components into a single data frame
regression_summaries <- bind_rows(
  summary_T1 %>% mutate(Day = "Day_T1", R_squared = glance_T1$r.squared),
  summary_T2 %>% mutate(Day = "Day_T2", R_squared = glance_T2$r.squared),
  summary_T3 %>% mutate(Day = "Day_T3", R_squared = glance_T3$r.squared),
  summary_T4 %>% mutate(Day = "Day_T4", R_squared = glance_T4$r.squared),
  summary_T5 %>% mutate(Day = "Day_T5", R_squared = glance_T5$r.squared)
) %>%
  filter(term == "(Intercept)" | term == "GPS_noerdlB") %>%
  select(Day, term, estimate, std.error, statistic, p.value,R_squared)

# Display the regression summaries using `DT::datatable`
datatable(regression_summaries, 
          caption = "Linear Regression Summaries for Each Fertilisation Stage",
          options = list(pageLength = 10, scrollX = TRUE))


```
### Summary statictics of the deviance of fertilisation dates
```{r}

# Calculate mean
mean_stats <- latitude_plot_data %>%
  summarise(across(c(Diff_T1_model, Diff_T2_VB_GL, Diff_T3_BBCH32, Diff_T4_BBCH37, Diff_T5_BBCH49), 
                   ~mean(., na.rm = TRUE))) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Mean")

# Calculate median
median_stats <- latitude_plot_data %>%
  summarise(across(c(Diff_T1_model, Diff_T2_VB_GL, Diff_T3_BBCH32, Diff_T4_BBCH37, Diff_T5_BBCH49), 
                   ~median(., na.rm = TRUE))) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Median")

# Calculate minimum
min_stats <- latitude_plot_data %>%
  summarise(across(c(Diff_T1_model, Diff_T2_VB_GL, Diff_T3_BBCH32, Diff_T4_BBCH37, Diff_T5_BBCH49), 
                   ~min(., na.rm = TRUE))) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Min")

# Calculate maximum
max_stats <- latitude_plot_data %>%
  summarise(across(c(Diff_T1_model, Diff_T2_VB_GL, Diff_T3_BBCH32, Diff_T4_BBCH37, Diff_T5_BBCH49), 
                   ~max(., na.rm = TRUE))) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Max")

# Calculate standard deviation
sd_stats <- latitude_plot_data %>%
  summarise(across(c(Diff_T1_model, Diff_T2_VB_GL, Diff_T3_BBCH32, Diff_T4_BBCH37, Diff_T5_BBCH49), 
                   ~sd(., na.rm = TRUE))) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "SD")

# Combine the summary statistics into a single data frame using `Reduce`
summary_stats <- Reduce(function(x, y) merge(x, y, by = "Variable"), 
                        list(mean_stats, median_stats, min_stats, max_stats, sd_stats))

# Display the summary statistics using `DT::datatable`
datatable(summary_stats, 
          caption = "Summary Statistics for Fertilisation Day Deviance",
          options = list(pageLength = 10, scrollX = TRUE))


```
### Violin plot 
```{r}
ggplot(spatial_plot_data, aes(x = Fertilization_Stage, y = Date_Value)) +
  geom_violin(aes(fill = Fertilization_Stage), alpha = 0.6, width = 0.8, show.legend = FALSE) +  # Violin plot for T1 to T5
  geom_jitter(color = "black", width = 0.2, height = 0, size = 2, alpha = 0.8) +  # Jittered ponits for place
  theme_minimal() +
  labs(title = "Actual Fertilization Date Distribution by Stage and Place",
       x = "Fertilization Stage",
       y = "Fertilization Date") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1),  # Adjust x-axis label position
        legend.position = "right")  # Keep legend on the right if additional legends exist


```
